---
title: "Tests_and_Hypotheses"
author: "Aleksei Zverev"
date: "2023-06-05"
output: html_document
---

```{r setup, include=TRUE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(fig.width = 10, fig.height = 6)
 
library('phyloseq')
library('tidyverse')
library('vegan')



set.seed(5678)
setwd('~/Praktikum/TyumGU_2023/')
ps <- readRDS("ps.no.organells.RData")

# Select only samples with more than 8k reads per sample
ps <- prune_samples(sample_sums(ps) > 8000, ps)
ps
```

## Alpha-diversity

```{r}
# rarefy to minimal depth
ps.raref <- rarefy_even_depth(ps)

plot_richness(ps.raref, x = "MainPlant", measures=c("Observed", "Simpson"), color = "Soil")
```

Our suggestion: Observed ASV in the DP soil is lower, than in the BS soil. Lets test it.

Make a table with desirable diversity index

```{r}
alpha_div <- function(ps, measures){
  ps <- prune_taxa(taxa_sums(ps) > 0, ps)
  obs_sim <- estimate_richness(ps, measures = measures)
  Soil <- ps@sam_data$Soil
  MainPlant <- ps@sam_data$MainPlant
  alpha <- cbind(obs_sim, Soil, MainPlant)
  return(alpha)
}

alpha <- alpha_div(ps.raref, c("Observed", "Simpson"))
alpha
```


```{r}
wilcox.test(Observed ~ Soil, data = alpha)
```
We are not really able to use two-way ANOVA (unweighted groups), but we'll try

```{r}
aov <- aov(Observed ~ Soil + MainPlant, data = alpha)
summary(aov)
# TukeyHSD(aov, conf.level=0.95)
```

No, there is no significant difference


## Beta-diversity

```{r}
ps.prop <- transform_sample_counts(ps, function(x) x/sum(x))
ord.pcoa.bray <- ordinate(ps.prop, method='PCoA', distance='bray')
plot_ordination(ps.prop, ord.pcoa.bray, color = 'MainPlant', shape = "Soil") +
    geom_point(size=3, alpha=0.7) + 
    theme_light()

```

Our suggestion: DP and ChZ forms separate clusters (Rye is in even more separate)

```{r}
dist <- phyloseq::distance(ps, method = "bray")
sample_df <- data.frame(sample_data(ps))
 
permanova <- adonis2(dist ~ Soil + MainPlant, data = sample_df)
permanova

# Use pairwise adonis as a post-hoc test
dp.only.ps <- subset_samples(ps, Soil == "DP")
dp.only.dist <- phyloseq::distance(dp.only.ps, method = "bray")
pairwiseAdonis::pairwise.adonis(dp.only.dist, dp.only.ps@sam_data$MainPlant)

chz.only.ps <- subset_samples(ps, Soil == "ChZ")
chz.only.dist <- phyloseq::distance(chz.only.ps, method = "bray")
pairwiseAdonis::pairwise.adonis(chz.only.dist, chz.only.ps@sam_data$MainPlant)
```

Yep, our suggestion has been proved!

## CCA

How about the connection between nutritional factors and microbial populations?

```{r}
# Collapse samples according to nutrition data. Add nutrition data to metadata section
ps@sam_data

ps@sam_data$Samples <- paste(ps@sam_data$Soil, ps@sam_data$MainPlant, ps@sam_data$Spot, sep=".")
ps.merged <- ps %>% merge_samples(group = "Samples")
ps.merged@sam_data


agro <- read.csv("agro.csv")
rownames(agro) <- agro$SampleID
agro

if (all(rownames(agro) == rownames(ps.merged@sam_data))) {
  ps.merged@sam_data <- sample_data(agro)
  print("Replaced")
}

ps.merged@sam_data
```

Make CCA for the top 1000 most abundant ASVs

```{r}
veganifyOTU <- function(physeq){
  require(phyloseq)
  if(taxa_are_rows(physeq)){physeq <- t(physeq)}
  return(as(otu_table(physeq), "matrix"))
}

ps.top1k <- names(sort(taxa_sums(ps.merged), decreasing = TRUE)[1:1000]) %>% 
  prune_taxa(ps.merged)

otus.ps.vegan <- veganifyOTU(ps.top1k)
metadata <- as(sample_data(ps.top1k), "data.frame")




vare.cca <- vegan::cca(otus.ps.vegan ~  pH + P + K + NH4, data=metadata)
vare.cca
```

Check the CCA model

```{r}
anova(vare.cca)
anova(vare.cca, by="terms")
vif.cca(vare.cca)
```
Our model is valid with four predictors: two significant and two unsignificant. I want to keep both.

```{r}
# uncomment this, if you want to see inside the model for the ...

# vare.cca$CCA$biplot # vectors
# vare.cca$CCA$u # samples
# vare.cca$CCA$v #ASVs
```
Ok, let's plot the data in two plots - for ASVs and for samples.

First, we have to fortify the model by taxonomic and sample information

```{r}
species.data <- vare.cca$CCA$v %>% 
               data.frame() %>% 
               mutate(ASV = rownames(.)) %>% 
               inner_join(data.frame(ASV = names(taxa_sums(ps.top1k)),
                                     Total.abund = taxa_sums(ps.top1k),
                                     ps.top1k@tax_table[,2], # Phylum
                                     ps.top1k@tax_table[,3]), # Class
                          by = "ASV")
species.data %>% head(10)


samples.data <- vare.cca$CCA$u %>% 
  data.frame() %>% 
  mutate(Names = rownames(.)) %>% 
  inner_join(ps@sam_data, by = c("Names" = "Samples"))

samples.data
```


```{r}
# plot species
ggplot() +
  geom_point(data=species.data,
             aes(x=CCA1, y=CCA2, color=Phylum, size=Total.abund), alpha=0.9) +
  geom_segment(data = vare.cca$CCA$biplot %>% data.frame(), 
               aes(x = 0, xend = CCA1, y = 0, yend = CCA2), 
               alpha=0.8, color = "black",arrow = arrow(angle = 3)) +
  geom_text(data = vare.cca$CCA$biplot %>% 
                    data.frame() %>% 
                    mutate(Label = rownames(.)), 
            aes(x=CCA1, y=CCA2, label= Label,
                hjust = -0.5), size=4) +
  theme_light() +
  ggtitle("A: Species")
```

Plot the same picture for different major (10 the most abundant) phyla

```{r}
major.phyla <- species.data %>% 
  group_by(Phylum) %>% 
  summarize(sum = sum(Total.abund)) %>% 
  arrange(desc(sum)) %>% 
  select(Phylum) %>% 
  head(10) %>% 
  as.vector()


for (i in major.phyla$Phylum) {
  p <- ggplot() +
    geom_point(data=species.data,
               aes(x=CCA1, y=CCA2, size=Total.abund), alpha=0.2, color="grey80") +
    geom_point(data=species.data %>% filter(Phylum == i),
               aes(x=CCA1, y=CCA2, color=Class, size=Total.abund), alpha=0.9) +
    geom_segment(data = vare.cca$CCA$biplot %>% data.frame(), 
                 aes(x = 0, xend = CCA1, y = 0, yend = CCA2), 
                 alpha=0.8, color = "black",arrow = arrow(angle = 3)) +
    geom_text(data = vare.cca$CCA$biplot %>% 
                      data.frame() %>% 
                      mutate(Label = rownames(.)), 
              aes(x=CCA1, y=CCA2, label= Label,
                  hjust = -0.5), size=4) +
    theme_light() +
    ggtitle(i)
  print(p)
}
```



```{r}
# plot samples
ggplot() +
  geom_point(data=samples.data,
             aes(x=CCA1, y=CCA2, color=MainPlant, shape = Soil), size=3, alpha=0.7) +
  geom_segment(data = vare.cca$CCA$biplot %>% data.frame(), 
               aes(x = 0, xend = CCA1, y = 0, yend = CCA2), 
               alpha=0.8, color = "black",arrow = arrow(angle = 3)) +
  geom_text(data = vare.cca$CCA$biplot %>% 
                    data.frame() %>% 
                    mutate(Label = rownames(.)), 
            aes(x=CCA1, y=CCA2, label= Label,
                hjust = -0.5), size=4) +
  theme_light() +
  ggtitle("B. Samples")
```

